Optimization model for multi-product multi-period multi-supplier raw-material selection and composition, and order quantity problem with minimum one-year order quantity contract

### 2023-03-09 Thesis/MethodX/Manuscript/Manuscript.Rmd
manuskrip final

```{r,include=FALSE}
setwd("~/209_ITB/Thesis/MethodX/Manuscript")

@@

This paper is written as follow. After this introduction, we describe the production planning and inventory control in the company in Section 2. After that we derive the optimization model in Section 3. In this section, we discuss the definition of the objective function which expresses all criteria for decision making. Then, in Section 5 we present the result of our numerical experimentation for obtaining the optimal solution. We end this paper by discussing the sensitivity analysis of the decision criteria to the optimal solution which we present in Section 5.

@@

df = read_excel("~/209_ITB/Thesis/Overleaf Version/data input.xlsx",sheet = "matriks produk x minggu")
colnames(df)[2:5] = paste0("week ",1:4)
df[1:11,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 1-11") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw -material demand (in kg)" = 4))
```

@@

```{r,echo=FALSE,message=FALSE,warning=FALSE}
df[12:51,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 12-51") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw -material demand (in kg)" = 4))
```

@@

>>>>

```{r setup,include=FALSE}
setwd("~/209_ITB/Thesis/MethodX/Manuscript")

@@

This paper is written as follow. After this introduction, we describe the production planning and inventory control in the company in Section 2. After that we derive the optimization model in Section 3. In this section, we discuss the definition of the objective function which expresses all criteria for decision making. Then, in Section 4 we present the result of our numerical experimentation for obtaining the optimal solution. We end this paper by discussing the sensitivity analysis of the decision criteria to the optimal solution which we present in Section 5.

@@

my_comma = scales::label_comma(accuracy = .1, big.mark = ",", decimal.mark = ".")

df = read_excel("~/209_ITB/Thesis/Overleaf Version/data input.xlsx",sheet = "matriks produk x minggu")
colnames(df)[2:5] = paste0("week ",1:4)

df$`week 1` = my_comma(df$`week 1`)
df$`week 2` = my_comma(df$`week 2`)
df$`week 3` = my_comma(df$`week 3`)
df$`week 4` = my_comma(df$`week 4`)

df[1:11,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 1-11") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw -material demand (in kg)" = 4))
```

@@

```{r,echo=FALSE,message=FALSE,warning=FALSE}
df[12:51,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 12-51") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw-material demand (in kg)" = 4))
```

@@

my_comma = scales::label_comma(accuracy = .1, big.mark = ",", decimal.mark = ".")

df = read_excel("parameter.xlsx")
df$`1` = my_comma(df$`1`)
df$`2` = my_comma(df$`2`)
df$`3` = my_comma(df$`3`)
df$`4` = my_comma(df$`4`)
df$`5` = my_comma(df$`5`)
df$`6` = my_comma(df$`6`)

df$code_gula[2] = '$\\sigma_k$ (in kg)'
df$code_gula[3] = '$z_{0k}$ (in kg)'
df$code_gula[4] = '$m_{0k}$ (in kg)'
df$code_gula[5] = '$ss$ (in kg)'

colnames(df)[1] = " "

df %>% 
  knitr::kable("latex", align="c", booktabs=TRUE, escape = F,
               caption = "the other parameter's values") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw-material" = 6)) %>% 
  footnote(general = c("maxcap (in kg) = 1,427,000"))
```

From Table 2 we know that raw material 4 must be purchased since item 49 and 50 can be produced just by using raw material 4.  The price of raw material 4 is the lowest one. But the minimal one-year order quantity is the second smallest. So that we may guess that on the optimal solution, $x_4$ will have a big value but is not the biggest one among others.

## Optimal solution

We solve the optimization problem by creating computer codes using R language (version 4) where the optimization problem formulation and the solution technique used are referred to dplyr [@dplyr] and ompr [@ompr] libraries. This program runs on a computer with the Linux Ubuntu 20 LTS operating system with an Intel i7 8 Cores processor and 16 GB RAM.

The values of $x_k,k \in \mathfrak{N},$ are given in the second column of Table 6 in the following. The weekly deliveries are given on columns 4 up to 6.  We see that the total one-month order quantity exceeds the minimum one-month order quantity $\sigma_k,$ so that set of Constraints I is satisfied.

df = data.frame(
  x1 = c(1:6,"Total"),
  x2 = c("192,649","543,385","90,742","56,508","52,958","202,853","1,139,094"),
  x3 = c("6,553","354,839","41,092","33,264","42,608","70,728","6,553"),
  x4 = c("65,295","23,595","26,550","209","10,350","111,075","65,295"),
  x5 = c("81,722","96,122","14,400","18,369","0","0","81,722"),
  x6 = c("39,079","68,829","8,700","4,666","0","21,050","39,079")
)

colnames(df)[1]   = "Raw Material"
colnames(df)[2]   = "Order Qty\n(in kg)"
colnames(df)[3]   = "Week 1"
colnames(df)[4]   = "Week 2"
colnames(df)[5]   = "Week 3"
colnames(df)[6]   = "Week 4"

df %>% 
  knitr::kable("latex",align="c", booktabs=TRUE, escape = F,
               caption = "Total order quantity and weekly deliveries") %>% 
  add_header_above(c(" " = 2,"weekly deliveries (in kg)" = 4)) %>% 
  kable_styling(latex_options = "hold_position")
```

Besides the values $x_k$ above, we also get $\hat{x}_{ijk}, a_{ijk}, b_{ijk}$ and values $z_{jk}$ in the optimal solution. We have checked that set of __Constraints II – V__ are satisfied by this optimal solution. For brevity, we do not present those values here. We just present and discuss some of them, to show that the remaining constraints are satisfied.

Most values of $b_{ijk}$ are $0.5$ which mean most of items are produced by using a composition of two raw materials. The values of $b_{ijk}$ may vary from one week to another wee, as illustrated in Table 7. The composition of raw materials for item 46 in week 1is different from the composition in week 2. Furthermore, Table 7 shows us that the proportions of raw materials used for producing an item are the same. We have checked this property through all  for all  and for all  so that we are sure that set of __Constraints VI__ are satisfied.

my_comma = scales::label_comma(accuracy = .01, big.mark = ",", decimal.mark = ".")

df = data.frame(
  item = c(rep(45,4),rep(46,4)),
  week = c(1:4,1:4),
  x1   = c(0,0.5,0.5,0.5,0.333,0.5,0,0),
  x2   = c(0.5,0.5,0.5,0.5,0.333,0.5,0,0),
  x3   = rep(0,8),
  x4   = c(0.5,0,0,0,0.333,0,0,0),
  x5   = rep(0,8),
  x6   = rep(0,8)
) %>% 
  mutate(x1 = my_comma(x1),
         x2 = my_comma(x2),
         x3 = my_comma(x3),
         x4 = my_comma(x4),
         x5 = my_comma(x5),
         x6 = my_comma(x6))

colnames(df)[3:8] = 1:6

df %>% 
  knitr::kable("latex", align="c", booktabs=TRUE, escape = F,
               caption = "the values of $b_{ijk}$ for $i = 45$ and $i = 46$") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 2,"raw-material" = 6)) 
```

The following table shows the composition of raw materials for items 49 and 50. We can see that these items are produced by using raw material 4, which confirm the flexibility matrix in Table 3.

my_comma = scales::label_comma(accuracy = 1, big.mark = ",", decimal.mark = ".")

df = data.frame(
  item = c(rep(49,4),rep(50,4)),
  week = c(1:4,1:4),
  x1   = rep(0,8),
  x2   = rep(0,8),
  x3   = rep(0,8),
  x4   = c(0,1,0,0,1,0,1,0),
  x5   = rep(0,8),
  x6   = rep(0,8)
) %>% 
  mutate(x1 = my_comma(x1),
         x2 = my_comma(x2),
         x3 = my_comma(x3),
         x4 = my_comma(x4),
         x5 = my_comma(x5),
         x6 = my_comma(x6))

colnames(df)[3:8] = 1:6

df %>% 
  knitr::kable("latex", align="c", booktabs=TRUE, escape = F,
               caption = "the values of $b_{ijk}$ for $i = 49$ and $i = 50$") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 2,"raw-material" = 6)) 
```

The fulfillment of the safety stock constraint and the maximum capacity constraint (set of __Constraints VII and VIII__) can be seen in the following table.

df = data.frame(
  raw = 1:6,
  x1  = c("2,501","181,743","2,500","2,500","2,500","2,500"),
  x2  = c("2,501","2,501","2,500","2,501","2,500","2,500"),
  x3  = c("2,500","2,500","2,500","2,501","2,500","2,500"),
  x4  = c(rep("2,500",6))
)

colnames(df)[1]   = "Raw material"
colnames(df)[2:5] = paste0("week ",1:4)

df %>% 
  knitr::kable("latex", align="c", booktabs=TRUE, escape = F,
               caption = "The fulfillment of the safety stock constraint and the maximum capacity constraint (set of Constraints VII and VIII)") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"end-of-week stock (in kg)" = 4)) 
```

# Conclusion and Discussion

In this paper we present mixed-integer linear program (MILP) for raw-material selection and composition, and order allocation problem faced by a beverage company. We performed a number of numerical experimentation to check the optimal solutions for various demands and for various objective functions.  From these experimentation we are sure that our MILP gives the desired optimal solutions.

It should be noticed that the definition of the objective function which accommodate the minimum one-year order quantity contract yields a balancing of purchase price criteria and the minimum one-year order quantity criteria. In the following table we represent the optimal solutions obtained by using different objective functions. Notice that the total raw materials purchased in all optimal solutions are the same, i.e. 1,139,094 kg.

Table 10 The different of the optimal solutions obtained by using different objective functions.

- Objective function 1:

$$
\frac{1}{2} hc \sum_{j \in M} \sum_{k \in \mathfrak{N}} (z_{(j-1)k} + z_{jk} + \hat{x}_{jk})
$$

- Objective function 2:

$$
\frac{1}{2} hc \sum_{j \in M} \sum_{k \in \mathfrak{N}} (z_{(j-1)k} + z_{jk} + \hat{x}_{jk}) + \sum_{k \in \mathfrak{N}} c_k x_k
$$

- Objective function 3:

$$
\frac{1}{2} hc \sum_{j \in M} \sum_{k \in \mathfrak{N}} (z_{(j-1)k} + z_{jk} + \hat{x}_{jk}) + \sum_{k \in \mathfrak{N}} c_k x_k - \sum_{k \in \mathfrak{N}} \alpha_k mo_k x_k
$$

df = data.frame(
  raw = c(1:6,"Total"),
  x1  = c("276,780","464,388","35,645","214,811","29,400","118,070","1,139,094"),
  x2  = c("363,486","532,404","20,402","60,124","16,888","145,791","1,139,094"),
  x3  = c("192,649","543,385","90,742","56,508","52,958","202,853","1,139,094")
)

colnames(df)[1]   = "Raw material"
colnames(df)[2:4] = paste0("obj function ",1:3)

df %>% 
  knitr::kable("latex", align="c", booktabs=TRUE, escape = F,
               caption = "Optimal solution comparison between different objective functions") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"Optimal solution" = 3)) 
```

From Table 10 we see if we just use the purchase price in the objective function, raw material 4 is purchased with the biggest amount. But since in the distribution of $mo_k$ raw material 4 has the second smallest value, then if we consider this distribution in the objective function, raw material 4 is purchased with a smaller amount.

The last thing to be discussed is the multiplier parameter $\alpha_k$. These parameters should be set as a positive number in early months of the year, when the minimum one-year order quantity contracts are still far away to be fulfilled. As soon as the contract for raw material $k$ is fulfilled, we can set $\alpha_{k=0}$.

# Acknowledment

The research in this paper is  partially funded by the Penelitian Tesis Magister research grant, Ministry of Research and Technology of the Republic of Indonesia.
